---
title: "GxE Analysis"
author: "Oluwafemi Oyedele"
date: "1/7/2022"
output: html_document
---

# Load required library for Single Trial Analysis

```{r}
#intstall.packages('statgenSTA')
#install.packages('statgenGxE')
#install.packages('here')

library(readxl)
library(statgenGxE)
library(statgenSTA)
library(tidyverse)
```

```{r}
data <- read_excel(here::here('Data/Copy of Sensory Data for Bayesian Analysis.xlsx'))
```

################################################# 

######## Analysis for Mealiness   #######################

################################################# 



# GGE Analysis
A Genotype plus Genotype by Environment (GGE) analysis is very similar to an AMMI analysis. The difference is in the first step where, instead of genotype and environment, only environment is fitted as a main effect in the model. Therefore, the principal component analysis is performed on the genotype main effect and the interaction jointly. In essence, the GGE analysis fits a principal components model with two components to the two-way genotype by environment table with data centered per environment.

# Objectives
# GxE Analysis
# Stability Analysis


# Create a TD object for GxE analysis for Mealiness

```{r}
dataTD <- createTD(data=data,genotype = 'Code',trial = 'Env')
```

## Perform a Finlay-Wilkinson analysis for all Trials

```{r}
Finlay1 <- gxeFw(TD=dataTD,trait='Mealiness')
```

## Create a scatter plot of Finlay Wilkins Analysis

```{r}
plot(Finlay1,plotType = 'scatter')
```

# Mixed Model Analysis for GxE Table of Means

```{r}
Mixed_VarComp <- gxeVarComp(TD = dataTD,trait = 'Mealiness')
summary(Mixed_VarComp)
```

# Extract Variance Component

```{r}
vc(Mixed_VarComp)
```

# Compute Heritabilty

```{r}
herit(Mixed_VarComp)
```

##### GxGGE analysis

```{r}
# Boxplot showing the distribution of the dataset

plot(dataTD, plotType = "box", traits = "Mealiness",orderBy = "descending")


GGE_Mealiness <- gxeGGE(TD=dataTD,trait='Mealiness')
summary(GGE_Mealiness)

# gxeAmmi() for AMMI ANALYSIS
```

# Visualize the results using a biplot

```{r}
plot(GGE_Mealiness,plotType = 'GGE2',scale = 0.5,sizeGeno = 3)+theme_classic()
plot(GGE_Mealiness,plotType = 'AMMI1',scale = 0.5,sizeGeno = 3)+theme_classic()+theme_classic()
plot(GGE_Mealiness,plotType = 'AMMI2',scale = 0.5,sizeGeno = 3)+theme_classic()
```

## Identifying mega environmenet

```{r}
Mega_Mealiness <- gxeMegaEnv(TD = dataTD,trait = 'Mealiness')
summary(Mega_Mealiness)
```

### Let me visualize the Mega_Environment

```{r}
plot(Mega_Mealiness)
```

## Stability Analysis

```{r}
Stability_Mealiness <- gxeStability(TD = dataTD,trait = 'Mealiness')
summary(Stability_Mealiness) # This shows top 10 performing genotype
#summary(Stability_Mealiness,pctGeno=2) # This shows top 2 performing genotype
```

# Create plot for different Stability measures

```{r}
plot(Stability_Mealiness)
```

################################################# 

##### Analysis for Fibre

###### 

################################################# 

## Perform a Finlay-Wilkinson analysis for all Trials

```{r}
Finlay2 <- gxeFw(TD=dataTD,trait='Fibre')
```

## Create a scatter plot of Finlay Wilkins Analysis

```{r}
plot(Finlay2,plotType = 'scatter')
```

# Mixed Model Analysis for GxE Table of Means

```{r}

# Boxplot showing the distribution of the dataset

plot(dataTD, plotType = "box", traits = "Fibre",orderBy = "descending")


Mixed_VarComp2 <- gxeVarComp(TD = dataTD,trait = 'Fibre')
summary(Mixed_VarComp2)
```

# Extract Variance Component

```{r}
vc(Mixed_VarComp2)
```

# Compute Heritabilty

```{r}
herit(Mixed_VarComp2)
```

##### GxGGE analysis

```{r}
GGE_Fibre2 <- gxeGGE(TD=dataTD,trait='Fibre')
summary(GGE_Fibre2)
```

# Visualize the results using a biplot

```{r}
plot(GGE_Fibre2,plotType = 'GGE2',scale = 0.5,sizeGeno = 3)+theme_classic()
plot(GGE_Fibre2,plotType = 'AMMI1',scale = 0.5,sizeGeno=3)+theme_classic()
plot(GGE_Fibre2,plotType = 'AMMI2',scale = 0.5,sizeGeno = 3)+theme_classic()
```

## Identifying mega environmenet

```{r}
Mega_Fibre2 <- gxeMegaEnv(TD = dataTD,trait = 'Fibre')
summary(Mega_Fibre2)
```

### Let me visualize the Mega_Environment

```{r}
plot(Mega_Fibre2)
```

## Stability Analysis

```{r}
Stability_Fibre2 <- gxeStability(TD = dataTD,trait = 'Fibre')
summary(Stability_Fibre2) # This shows top 10 performing genotype
#summary(Stability_Mealiness,pctGeno=2) # This shows top 2 performing genotype
```

# Create plot for different Stability measures

```{r}
plot(Stability_Fibre2)
```

##### Analysis for ADH

###### 

################################################# 

## Perform a Finlay-Wilkinson analysis for all Trials

```{r}
Finlay3 <- gxeFw(TD=dataTD,trait='ADH')
```

## Create a scatter plot of Finlay Wilkins Analysis

```{r}
plot(Finlay3,plotType = 'scatter')
```

# Mixed Model Analysis for GxE Table of Means

```{r}
# Boxplot showing the distribution of the dataset

plot(dataTD, plotType = "box", traits = "ADH",orderBy = "descending")


Mixed_VarComp3 <- gxeVarComp(TD = dataTD,trait = 'ADH')
summary(Mixed_VarComp3)
```

# Extract Variance Component

```{r}
vc(Mixed_VarComp3)
```

# Compute Heritabilty

```{r}
herit(Mixed_VarComp3)
```

##### GxGGE analysis

```{r}
GGE_ADH3 <- gxeGGE(TD=dataTD,trait='ADH')
summary(GGE_ADH3)
```

# Visualize the results using a biplot

```{r}
plot(GGE_ADH3,plotType = 'GGE2',scale = 0.5,sizeGeno = 3)+theme_classic()
plot(GGE_ADH3,plotType = 'AMMI1',scale = 0.5,sizeGeno=3)+theme_classic()
plot(GGE_ADH3,plotType = 'AMMI2',scale = 0.5,sizeGeno=3)+theme_classic()
```

## Identifying mega environmenet

```{r}
Mega_ADH3 <- gxeMegaEnv(TD = dataTD,trait = 'ADH')
summary(Mega_ADH3)
```

### Let me visualize the Mega_Environment

```{r}
plot(Mega_ADH3)
```

## Stability Analysis

```{r}
Stability_ADH3 <- gxeStability(TD = dataTD,trait = 'ADH')
summary(Stability_ADH3) # This shows top 10 performing genotype
#summary(Stability_Mealiness,pctGeno=2) # This shows top 2 performing genotype
```

# Create plot for different Stability measures

```{r}
plot(Stability_ADH3)
```

################################################# 

##### Analysis for Softness

###### 

################################################# 

## Perform a Finlay-Wilkinson analysis for all Trials

```{r}
Finlay4 <- gxeFw(TD=dataTD,trait='Softness')
```

## Create a scatter plot of Finlay Wilkins Analysis

```{r}
plot(Finlay4,plotType = 'scatter')
```

# Mixed Model Analysis for GxE Table of Means

```{r}
# Boxplot showing the distribution of the dataset

plot(dataTD, plotType = "box", traits = "Softness",orderBy = "descending")


Mixed_VarComp4 <- gxeVarComp(TD = dataTD,trait = 'Softness')
summary(Mixed_VarComp4)
```

# Extract Variance Component

```{r}
vc(Mixed_VarComp4)
```

# Compute Heritabilty

```{r}
herit(Mixed_VarComp4)
```

##### GxGGE analysis

```{r}
GGE_Softness4 <- gxeGGE(TD=dataTD,trait='Softness')
summary(GGE_Softness4)
```

# Visualize the results using a biplot

```{r}
plot(GGE_Softness4,plotType = 'GGE2',scale = 0.5,sizeGeno = 3)+theme_classic()
plot(GGE_Softness4,plotType = 'AMMI1',scale = 0.5,sizeGeno=3)+theme_classic()
plot(GGE_Softness4,plotType = 'AMMI2',scale = 0.5,sizeGeno=3)+theme_classic()


```

## Identifying mega environmenet

```{r}
Mega_Softness4 <- gxeMegaEnv(TD = dataTD,trait = 'Softness')
summary(Mega_Softness4)
```

### Let me visualize the Mega_Environment

```{r}
plot(Mega_Softness4)
```

## Stability Analysis

```{r}
Stability_Softness4 <- gxeStability(TD = dataTD,trait = 'Softness')
summary(Stability_Softness4) # This shows top 10 performing genotype
#summary(Stability_Mealiness,pctGeno=2) # This shows top 2 performing genotype
```

# Create plot for different Stability measures

```{r}
plot(Stability_Softness4)
```

################################################# 

##### Analysis for Taste

###### 

################################################# 

## Perform a Finlay-Wilkinson analysis for all Trials

```{r}
Finlay5 <- gxeFw(TD=dataTD,trait='Taste')
```

## Create a scatter plot of Finlay Wilkins Analysis

```{r}
plot(Finlay5,plotType = 'scatter')
```

# Mixed Model Analysis for GxE Table of Means

```{r}
# Boxplot showing the distribution of the dataset

plot(dataTD, plotType = "box", traits = "Taste",orderBy = "descending")


Mixed_VarComp5 <- gxeVarComp(TD = dataTD,trait = 'Taste')
summary(Mixed_VarComp5)
```

# Extract Variance Component

```{r}
vc(Mixed_VarComp5)
```

# Compute Heritabilty

```{r}
herit(Mixed_VarComp5)
```

##### GxGGE analysis

```{r}
GGE_Taste5 <- gxeGGE(TD=dataTD,trait='Taste')
summary(GGE_Taste5)
```

# Visualize the results using a biplot

```{r}
plot(GGE_Taste5,plotType = 'GGE2',scale = 0.5,sizeGeno = 3)+theme_classic()
plot(GGE_Taste5,plotType = 'AMMI1',scale = 0.5,sizeGeno=3)+theme_classic()
plot(GGE_Taste5,plotType = 'AMMI2',scale = 0.5,sizeGeno=3)+theme_classic()
```

## Identifying mega environmenet

```{r}
Mega_Taste5 <- gxeMegaEnv(TD = dataTD,trait = 'Taste')
summary(Mega_Taste5)
```

### Let me visualize the Mega_Environment

```{r}
plot(Mega_Taste5)
```

## Stability Analysis

```{r}
Stability_Taste5 <- gxeStability(TD = dataTD,trait = 'Taste')
summary(Stability_Taste5) # This shows top 10 performing genotype
#summary(Stability_Mealiness,pctGeno=2) # This shows top 2 performing genotype
```

# Create plot for different Stability measures

```{r}
plot(Stability_Taste5)
```

################################################# 

##### Analysis for Colour

###### 

################################################# 

## Perform a Finlay-Wilkinson analysis for all Trials

```{r}
Finlay6 <- gxeFw(TD=dataTD,trait='Colour')
```

## Create a scatter plot of Finlay Wilkins Analysis

```{r}
plot(Finlay6,plotType = 'scatter')
```

# Mixed Model Analysis for GxE Table of Means

```{r}
Mixed_VarComp6 <- gxeVarComp(TD = dataTD,trait = 'Colour')
summary(Mixed_VarComp6)
```

# Extract Variance Component

```{r}
vc(Mixed_VarComp6)
```

# Compute Heritabilty

```{r}
herit(Mixed_VarComp6)
```

##### GxGGE analysis

```{r}
# Boxplot showing the distribution of the dataset

plot(dataTD, plotType = "box", traits = "Colour",orderBy = "descending")


GGE_Colour6 <- gxeGGE(TD=dataTD,trait='Colour')
summary(GGE_Colour6)
```

# Visualize the results using a biplot

```{r}
plot(GGE_Colour6,plotType = 'GGE2',scale = 0.5,sizeGeno = 3)+theme_classic()
plot(GGE_Colour6,plotType = 'AMMI1',scale = 0.5,sizeGeno=3)+theme_classic()
plot(GGE_Colour6,plotType = 'AMMI2',scale = 0.5,sizeGeno=3)+theme_classic()
```

## Identifying mega environmenet

```{r}
Mega_Colour6 <- gxeMegaEnv(TD = dataTD,trait = 'Colour')
summary(Mega_Colour6)
```

### Let me visualize the Mega_Environment

```{r}
plot(Mega_Colour6)
```

## Stability Analysis

```{r}
Stability_Colour6 <- gxeStability(TD = dataTD,trait = 'Colour')
summary(Stability_Colour6) # This shows top 10 performing genotype
#summary(Stability_Mealiness,pctGeno=2) # This shows top 2 performing genotype
```

# Create plot for different Stability measures

```{r}
plot(Stability_Colour6)
```
